---
title: "p8105_hw6_ar4459_practice"
author: "ASHLEY ROMO"
date: "2023-11-30"
output: github_document
---

load key packages
```{r}
library(tidyverse)
library(p8105.datasets)
library(modelr)

set.seed(1234)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


### Problem 2
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

First, I created a bootstrap function. 
```{r}
# bootstrap function 
boot_sample = function(df) {
  
  sample_frac(df, replace = TRUE)
}
```

Next, I apply the bootsrap function and use broom::glance to calculate the r square value.
```{r}
# apply bootstrap for r square
boot_results_r =
  tibble(strap_number = 1:50) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(weather_df)),
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::glance))|> 
  select(-strap_sample, -models) |> 
    unnest(results) |> 
   janitor::clean_names() 
  
#r.square plot
boot_results_r |> 
  ggplot(aes(x = r_squared)) +
  geom_density() +
  labs(
    x = "R Squared",
    y = "Density",
    title = "Distribution of R Squared"
  )
```
The r square plot is unimodal. The max r squared is `r max(boot_results_r |> pull(r_squared))`.


Now, I apply the bootstrap function to the data frame and use the broom::tidy function to calculate the beta estimates. I use pivot_wider to isolate the values for tmin and prcp and also rename tmin to beta1 and prcp to beta2. I use mutate to calculate the log of beta1 times beta2. Lastly, I use ggplot to plot the distribution of log(beta1*beta2).
```{r}
# apply bootstrap for log(b1*b2)
boot_results_beta =
  tibble(strap_number = 1:50) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(weather_df)),
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::tidy))|> 
    unnest(results) |> 
  janitor::clean_names() |> 
  select(strap_number, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate) |> 
  rename(
    beta1 = tmin,
    beta2 = prcp) |> 
  mutate(
    log_betas = log(beta1*beta2)) |> 
  drop_na()

#plot for log betas
boot_results_beta |> 
  ggplot(aes(x = log_betas)) +
  geom_density() +
  labs(
    x = "log(beta1*beta2)",
    y = "density",
    title = "Distribution of log of beta1*beta2")
```


Using the quintile function, I calculate the 95% confidence interval for the log(beta1*beta2) and the r squared.
```{r}
#95% confidence interval for r square
r_ci =
  boot_results_r |> 
  summarize(
    ci_lower_est = quantile(r_squared, 0.025),
    ci_upper_est = quantile(r_squared, 0.0975)) |> 
  knitr::kable(digits = 2)

#95% confidence interval for log(beta1*beta2)
log_ci =
  boot_results_beta |> 
  summarize(
    ci_lower_est = quantile(log_betas, 0.025),
    ci_upper_est = quantile(log_betas, 0.0975)) |> 
  knitr::kable(digits = 2)
```


### Problem 3


First, I load the data and drop any missing values. 
```{r}
#data cleaning
bw_data = 
  read_csv("data/birthweight.csv") |> 
  drop_na()
```

Next, I propose a linear model. I chose to include variables that are associated with birthweight. These variables are delwt, gaweeks, pnumlbw, ppwt, and wtgain.
```{r}
#propose a linear regression model
linear_mod = lm(bwt ~ delwt + gaweeks + pnumlbw + ppwt + wtgain , data = bw_data) 

```

Next, I use modelr::add_predictions and modelr::add_residuals to calculate and then plot the predicted values against the residual values
```{r}
#plot predictions (fitted) and residual values 
bw_data  |> 
  modelr::add_predictions(linear_mod) |>
  modelr::add_residuals(linear_mod) |> 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() 
```

Here, I listed the two models according to the homework description.
```{r}
#model 1
model_one = lm(bwt ~ blength + gaweeks, data = bw_data)

#model 2
model_two = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = bw_data)
```

Cross Validation 
```{r}
#use crosv_mc to do the crossvaludation on the train and test 
cv_df = 
  bw_data |> 
  crossv_mc(n = 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

# cross validation on three different models
cv_results =
  cv_df|> 
  mutate(
    linear_mod = map(train, ~lm(bwt ~ delwt + gaweeks + pnumlbw + ppwt + wtgain, data = .x)),
    model_one = map(train, ~lm(bwt ~ blength + gaweeks,  data = .x)),
    model_two = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = .x))) |> 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, \(mod, df) rmse(mod, df)),
    rmse_model_one = map2_dbl(model_one, test, \(mod, df) rmse(mod, df)),
    rmse_model_two = map2_dbl(model_two, test, \(mod, df) rmse(mod, df))
  ) 

cv_results_rmse =
  cv_results |> 
  select(.id, rmse_linear, rmse_model_one, rmse_model_two) |> 
  pivot_longer(
    rmse_linear:rmse_model_two,
    names_to = "model",
    names_prefix = "rmse_",
    values_to = "rmse") 

# plot of rmse values for linear model, model one, and model two
cv_results_rmse |> 
  group_by(model) |> 
  ggplot(aes(x = model, y = rmse)) +
  geom_violin()
```

